---
title: "Honey bee DMTS vibration study - statistical analysis and figures"
author: "Leslie Ng on behalf of all authors"
date: "2022-10-05"
output:
  html_document: null
  toc: yes
  pdf_document: default
toc_float: yes
toc_depth: 3
highlight: tango
theme: paper
self_contained: yes
mode: selfcontained
editor_options:
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../Output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this document we tested if free-moving honey bees could be trained using appetite-aversive conditioning to use the presence/absence of a vibration cue or the duration of a vibration cue to solve a symbolic delayed matching-to-sample task. This document provides details on data processing, statistical analysis and figures of the original manuscript submitted for peer review (titled: "Can honey bees use vibration cues to inform visual decisions in a symbolic delayed matching-to-sample task?").

The following code is written in the R programming language.

# Libraries 

### *Install packages*
```{r eval=FALSE}
install.packages("lme4")  # For fitting models
install.packages("ggplot2") # For plotting
```

### *Load packages*

```{r, message=FALSE, warning=FALSE}
library(lme4) # For fitting models
library(ggplot2) # For plotting
```



# Data importing

```{r, message=FALSE, warning=FALSE}

VIBDATA <- read.csv(
  "../Data/VIBDATA.csv", header=TRUE)

#Subset data

VIBDATA$BEEID <- as.factor(VIBDATA$BEEID) # Treat Bee ID as factor
CONTROL <- subset(VIBDATA, CONDITION == "CONTROL") # Control condition
SDISCRETE <- subset(VIBDATA, CONDITION == "SPEAKER_DISCRETE") # 250 Hz 0 vs 3 seconds condition
SCONT <- subset(VIBDATA, CONDITION == "SPEAKER_CONT") # 250 Hz 1 vs 3 seconds condition
MDISCRETE <- subset(VIBDATA, CONDITION == "MOTOR_DISCRETE") #68 Hz 0 vs 3 seconds condition

```



# Generalized linear models

```{r message=FALSE}

# Exp 1: 250 Hz 0 vs 3 seconds condition

lm1 <- glm(CHOICE~CNUM, family = binomial, data = MDISCRETE) 
summary(lm1)

# Exp 2: 250 Hz 1 vs 3 seconds condition

lm2 <- glm(CHOICE~CNUM, family = binomial, data = SDISCRETE) 
summary(lm2)

# Exp 3: 68 Hz 0 vs 3 seconds condition

lm3 <- glm(CHOICE~CNUM, family = binomial, data = SCONT) 
summary(lm3)

# Exp 4: Control condition

lm4 <- glm(CHOICE~CNUM, family = binomial, data = CONTROL) 
summary(lm4)
```



# Generate Figure 2 

### *Calculate predicted data points using lm1, lm2, lm3, and lm4*
```{r}

##### Generate predicted data points for Exp 1: 68 Hz 0 vs 3 seconds ####

choicenumlm  <- data.frame (CHOICENUM = seq(1,100, length.out = 100)) # Create dataframe from 1 to 100
Xgc1 <- model.matrix(~CHOICENUM, data = choicenumlm) # Create matrix of choice number
coeff <- c(0.019851, 0.001034) # Create vector of model coefficients pulled from lm1
eta <- Xgc1 %*% coeff # Calculate predicted mean proportion of correct choices
SE <- sqrt(diag(Xgc1 %*% vcov(lm1) %*% t(Xgc1))) # Calculate standard errors
  
  
PREDLM <- data.frame (CNUM = choicenumlm$CHOICENUM, ETA  = eta, STDERR = SE) # Create and add to a new dataframe
PREDLM$CONFINTLOW <-(PREDLM$ETA - 1.96*PREDLM$STDERR)  # Calculate confint lower bound
PREDLM$CONFINTHIGH <-(PREDLM$ETA + 1.96*PREDLM$STDERR) # Calculate confint upper bound
PREDLM$EXPETA <- (exp(PREDLM$ETA)/(1 + exp(PREDLM$ETA))) # Reverse the link function of the predicted mean proportion of correct choices
PREDLM$EXPCONFINTLOW <- (exp(PREDLM$CONFINTLOW)/(1 + exp(PREDLM$CONFINTLOW))) # Reverse the link function of the lower bound
PREDLM$EXPCONFINTHIGH <- (exp(PREDLM$CONFINTHIGH)/(1 +exp(PREDLM$CONFINTHIGH))) # Reverse the link function of the upper bound



#### Generate predicted data points for Exp 2: 250 Hz 0 vs 3 seconds ####
  
coeff2 <- c(5.739e-02, -8.007e-05) # Create vector of model coefficients pulled from lm2
eta2 <- Xgc1 %*% coeff2 # Calculate predicted mean proportion of correct choices
SE2 <- sqrt(diag(Xgc1 %*% vcov(lm2) %*% t(Xgc1))) # Calculate standard errors



PREDLM$ETA2 <- eta2 # Add mean proportion of correct choices to existing dataframe
PREDLM$STDERR2 <- SE2 # Add standard error to existing dataframe
PREDLM$CONFINTLOW2 <- (PREDLM$ETA2 - 1.96*PREDLM$STDERR2) # Calculate confint lower bound 
PREDLM$CONFINTHIGH2 <- (PREDLM$ETA2 + 1.96*PREDLM$STDERR2) # Calculate confint upper bound
PREDLM$EXPETA2 <- (exp(PREDLM$ETA2)/(1 + exp(PREDLM$ETA2))) # Reverse the link function of the predicted mean proportion of correct choices
PREDLM$EXPCONFINTLOW2 <- (exp(PREDLM$CONFINTLOW2)/(1 + exp(PREDLM$CONFINTLOW2))) # Reverse the link function of the lower bound
PREDLM$EXPCONFINTHIGH2 <- (exp(PREDLM$CONFINTHIGH2)/(1 +exp(PREDLM$CONFINTHIGH2))) # Reverse the link function of the upper bound
  


#### Generate predicted data points for Exp 3: 250 Hz 1 vs 3 seconds ####  

coeff3 <- c(0.0004838, 0.0003865) # Create vector of model coefficients pulled from lm3
eta3 <- Xgc1 %*% coeff3 # Calculate predicted mean proportion of correct choices
SE3 <- sqrt(diag(Xgc1 %*% vcov(lm3) %*% t(Xgc1))) # Calculate standard errors
  
  

PREDLM$ETA3 <- eta3 # Add mean proportion of correct choices to existing dataframe
PREDLM$STDERR3 <- SE3 # Add standard error to existing dataframe
PREDLM$CONFINTLOW3 <- (PREDLM$ETA3 - 1.96*PREDLM$STDERR3) # Calculate confint lower bound 
PREDLM$CONFINTHIGH3 <- (PREDLM$ETA3 + 1.96*PREDLM$STDERR3) # Calculate confint upper bound
PREDLM$EXPETA3 <- (exp(PREDLM$ETA3)/(1 + exp(PREDLM$ETA3))) # Reverse the link function of the predicted mean proportion of correct choices
PREDLM$EXPCONFINTLOW3 <- (exp(PREDLM$CONFINTLOW3)/(1 + exp(PREDLM$CONFINTLOW3))) # Reverse the link function of the lower bound 
PREDLM$EXPCONFINTHIGH3 <- (exp(PREDLM$CONFINTHIGH3)/(1 +exp(PREDLM$CONFINTHIGH3))) # Reverse the link function of the upper bound 
  
  
#### Generate predicted data points for Exp 4: colour control ####  
  
choicenumlm2  <- data.frame (CHOICENUM2 = seq(1,60, length.out = 60)) # Create new dataframe from 1 to 60
Xgc2 <- model.matrix(~CHOICENUM2, data = choicenumlm2) # Create new matrix of choice number
coeff4 <- c(-0.187623, 0.018451) # Create vector of model coefficients pulled from lm4
eta4 <- Xgc2 %*% coeff4 # Calculate predicted mean proportion of correct choices
SE4 <- sqrt(diag(Xgc2 %*% vcov(lm4) %*% t(Xgc2))) # Calculate standard errors 
  


PREDLM2 <- data.frame (CNUM = choicenumlm2$CHOICENUM2, ETA4  = eta4, STDERR4 = SE4) # Create and add to a new dataframe
PREDLM2$ETA4 <- eta4 # Add mean proportion of correct choices to new dataframe
PREDLM2$STDERR4 <- SE4 # Add standard error to new dataframe
PREDLM2$CONFINTLOW4 <- (PREDLM2$ETA4 - 1.96*PREDLM2$STDERR4) # Calculate confint lower bound
PREDLM2$CONFINTHIGH4 <- (PREDLM2$ETA4 + 1.96*PREDLM2$STDERR4) # Calculate confint upper bound
PREDLM2$EXPETA4 <- (exp(PREDLM2$ETA4)/(1 + exp(PREDLM2$ETA4))) # Reverse the link function of the predicted mean proportion of correct choices
PREDLM2$EXPCONFINTLOW4 <- (exp(PREDLM2$CONFINTLOW4)/(1 +exp(PREDLM2$CONFINTLOW4))) # Reverse the link function of the lower bound 
PREDLM2$EXPCONFINTHIGH4<- (exp(PREDLM2$CONFINTHIGH4)/(1+exp(PREDLM2$CONFINTHIGH4))) # Reverse the link function of the upper bound
  
```

## *Plot predicted values using ggplot (Figure 2)*

```{r, fig.height = 6, fig.width = 10, fig.align = "center"} 

ggplot() +
    geom_line(aes(x = CNUM, y = EXPETA, linetype = "dotted"), size = 0.5, data = PREDLM) +
    geom_line(aes(x = CNUM, y = EXPETA2, linetype = "dashed"), size = 0.5, data = PREDLM) +
    geom_line(aes(x = CNUM, y = EXPETA3, linetype = "twodash"), size = 0.5, data = PREDLM) +
    geom_line(aes(x = CNUM, y = EXPETA4, linetype = "solid"), size = 0.5, data = PREDLM2) +
    geom_ribbon(aes(x = CNUM, ymin = EXPCONFINTLOW4, ymax = EXPCONFINTHIGH4), 
                fill = "blue", alpha = 0.3, data = PREDLM2)  + # confint band
    geom_ribbon(aes(x = CNUM, ymin = EXPCONFINTLOW, ymax = EXPCONFINTHIGH), 
                fill = "purple", alpha = 0.3, data = PREDLM)  +
    geom_ribbon(aes(x = CNUM, ymin = EXPCONFINTLOW2, ymax = EXPCONFINTHIGH2), 
                fill = "red", alpha = 0.3, data = PREDLM)  +
    geom_ribbon(aes(x = CNUM, ymin = EXPCONFINTLOW3, ymax = EXPCONFINTHIGH3), 
                fill = "green", alpha = 0.3, data = PREDLM)  + 
    theme(panel.grid.major = element_line(linetype = "blank"), 
          panel.grid.minor = element_line(linetype = "blank"), 
          panel.background = element_rect(fill = "gray96"))  + theme(axis.title = element_text(size = 18)) +
    labs(x = "Choice number", y = "Mean proportion of correct choices (%)")+
    theme(axis.text = element_text(size = 16), legend.position="bottom", legend.title = element_blank(), legend.text = element_text(size = 16), axis.title.y=element_text(margin=margin(0,20,0,0)), axis.title.x=element_text(margin=margin(20,0,0,0))) +
scale_linetype_identity(guide = "legend", 
                        breaks = c("dotted", "dashed", "twodash", "solid"), 
                        labels = c("Exp 1: 68 Hz 0v3", "Exp 2: 250 Hz 0v3", "Exp 3: 250 Hz 1v3", "Exp 4: control")) 

    
```


